Positional surprisal (1-3grams) in CHILDES corpora

### 2019-09-12 Paper/paper.Rmd
bootstrap childes surprisal results

```{r setup, include=FALSE}
library(tidyverse)
library(ggdendro)
library(here)
library(feather)
library(ggdendro)

@@

```{r load data, include = F}
childes_results <- read_feather(here("../surprisal/Data/childes_words.feather"))
wiki_results <- read_feather(here("../surprisal/Data/all_spaces.feather"))
switchboard <- read_feather(here("../surprisal/Data/switchboard_words.feather"))
wiki_curves <- read_csv(here("../surprisal/Data/relative_ngrams.csv"))
```

@@

We have found a unique average distribution of information that appears to characterize the English language as a whole, regardless of medium (spoken or written) or utterance length. This distribution indicates that in English, the words we speak or write at the beginnings of utterances has little information, while the words we speak or write at the ends of utterances has a lot of information. The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from [@genzel2002entropy].

@@

```{r switchboard & bnc plot}

```

@@

```{r, echo=F, fig.height =2, fig.cap="Providence frequency-based information curves"}
childes_results %>% 
  filter(Language == "English", context == 0, length %in% c(5, 7, 9)) %>%
  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
    geom_point() + 
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:9) + 
    geom_line() + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

```{r, echo=F, fig.height =2, fig.cap="Providence bigram context-based information curves"}
childes_results %>% 
  filter(Language == "English", context == 1, length %in% c(5, 7, 9)) %>%
  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
    geom_point() + 
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:9) + 
    geom_line() + 
    xlab("Word position") + 
    ylab("Mean information")
```

```{r, echo=F, fig.height =2, fig.cap="Providence trigram context-based information curves"}
childes_results %>% 
  filter(Language == "English", context == 2, length %in% c(5, 7, 9)) %>%
  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
    geom_point() + 
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:9) + 
    geom_line() + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

```{r, echo=F, fig.height =3, fig.cap="CHILDES frequency-based information curves"}
childes_results %>% 
  filter(context == 0, length == 7) %>%
  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
    geom_point() + 
    facet_wrap(~Language) + 
    scale_x_continuous(breaks = 1:9) + 
    geom_line() + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

```{r, echo=F, fig.height =3, fig.cap="CHILDES bigram context-based information curves"}
childes_results %>% 
  filter(context == 1, length == 7) %>%
  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
    geom_point() + 
    facet_wrap(~Language) + 
    scale_x_continuous(breaks = 1:9) + 
    geom_line() + 
    xlab("Word position") + 
    ylab("Mean information")
```

```{r, echo=F, fig.height =3, fig.cap="CHILDES trigram context-based information curves"}
childes_results %>% 
  filter(context == 2, length == 7) %>%
  ggplot(aes(x = word_order, y = s, color = speaker_role)) + 
    geom_point() + 
    facet_wrap(~Language) + 
    scale_x_continuous(breaks = 1:9) + 
    geom_line() + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

Some lr results go here.

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(ggdendro)
library(here)
library(feather)
library(ggdendro)
library(gridExtra)

@@

```{r load data, include = F}
childes_results <- read_feather(here("Data/all_childes_bootstrap.feather"))
wiki_results <- read_feather(here("../surprisal/Data/all_spaces.feather"))
switch_results <- read_feather(here("Data/switch_bootstrap.feather"))
wiki_curves <- read_csv(here("../surprisal/Data/relative_ngrams.csv"))
wiki_lr <- read_feather(here("../surprisal/Wikipedia/models.feather"))
```

@@

```{r, echo=F, fig.height=2, fig.cap="Switchboard frequency-based information curves"}
switch_results %>% 
  filter(context == 0) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:11) + 
    xlab("Word position") + 
    ylab("Mean information")
```

We have found a unique average distribution of information that appears to characterize the English language as a whole, regardless of medium (spoken or written) or utterance length. This distribution indicates that in English, the words we speak or write at the beginnings of utterances have little information, while the words we speak or write at the ends of utterances have a lot of information. The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from [@genzel2002entropy].

@@

```{r, echo=F,fig.height=2,fig.cap="Switchboard bigram context-based information curves"}
switch_results %>% 
  filter(context == 1) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:11) + 
    xlab("Word position") + 
    ylab("Mean information")
```

```{r, echo=F,fig.height=2,fig.cap="Switchboard trigram context-based information curves"}
switch_results %>% 
  filter(context == 2) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:11) + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

```{r, echo=F, fig.height =2, fig.cap="Providence frequency-based information curves"}
childes_results %>% 
  filter(language == "English", context == 0) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
             color = speaker)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:9) + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

```{r, echo=F, fig.height =2, fig.cap="Providence bigram context-based information curves"}
childes_results %>% 
  filter(language == "English", context == 1) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
             color = speaker)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:9) + 
    xlab("Word position") + 
    ylab("Mean information")
```

```{r, echo=F, fig.height =2, fig.cap="Providence trigram context-based information curves"}
childes_results %>% 
  filter(language == "English", context == 2) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
             color = speaker)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~length) + 
    scale_x_continuous(breaks = 1:9) + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

```{r, echo=F, fig.height =3, fig.cap="CHILDES frequency-based information curves"}
childes_results %>% 
  filter(context == 0, length == 7) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
             color = speaker)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~language) + 
    scale_x_continuous(breaks = 1:7) + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

```{r, echo=F, fig.height =3, fig.cap="CHILDES bigram context-based information curves"}
childes_results %>% 
  filter(context == 1, length == 7) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
             color = speaker)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~language) + 
    scale_x_continuous(breaks = 1:7) + 
    xlab("Word position") + 
    ylab("Mean information")
```

```{r, echo=F, fig.height =3, fig.cap="CHILDES trigram context-based information curves"}
childes_results %>% 
  filter(context == 2, length == 7) %>%
  ggplot(aes(x = word_order, y = empirical_stat, ymin = ci_lower, ymax = ci_upper, 
             color = speaker)) + 
    geom_point() + 
    geom_pointrange() + 
    geom_line() +
    facet_wrap(~language) + 
    scale_x_continuous(breaks = 1:7) + 
    xlab("Word position") + 
    ylab("Mean information")
```

@@

```{r, echo=F, fig.cap="Some linear model results from Wikipedia and WALS features"}
sf <- wiki_lr %>% 
  filter(p.value == 0, estimate > 0) %>% 
  select(name) %>% 
  rename("Some Important Features" = name) %>% 
  slice(c(1, 10, 12, 15, 23, 29))

sf$Examples <- c("Noun-RC; RC-noun", 
                 "No cases; 6 cases", 
                 "SOV; VSO", 
                 "Only singular; sing. and plural", 
                 "Affix; distinct word", 
                 "Prefixes; suffixes")

sf %>% grid.table()
```

### 2019-09-12 Paper/paper.tex
bootstrap childes surprisal results

  \title{Speakers of diverse languages structure their utterances for efficient communication}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Josef Klafka and Daniel Yurovsky}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{2019-09-11}

@@

We have found a unique average distribution of information that appears to characterize the English language as a whole, regardless of medium (spoken or written) or utterance length. This distribution indicates that in English, the words we speak or write at the beginnings of utterances has little information, while the words we speak or write at the ends of utterances has a lot of information. The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from (Genzel \& Charniak, 2002).

@@

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}Providence frequency-based information curves}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{\label{fig:unnamed-chunk-2}Providence bigram context-based information curves}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{\label{fig:unnamed-chunk-3}Providence trigram context-based information curves}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{\label{fig:unnamed-chunk-4}CHILDES frequency-based information curves}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{\label{fig:unnamed-chunk-5}CHILDES bigram context-based information curves}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{\label{fig:unnamed-chunk-6}CHILDES trigram context-based information curves}
\end{figure}

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
\caption{\label{fig:unnamed-chunk-8}Some trigram information curves from the Wikipedia data}
\end{figure}

@@

Some lr results go here.

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1.pdf}
\caption{\label{fig:unnamed-chunk-9}wals features vs cosine similarity}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1.pdf}
\caption{\label{fig:unnamed-chunk-10}ldn features vs cosine similarity}
\end{figure}

>>>>

  \title{Speakers of diverse languages structure their utterances for efficient communication}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Josef Klafka and Daniel Yurovsky}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
      \predate{\centering\large\emph}
  \postdate{\par}
    \date{2019-09-12}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-1-1.pdf}
\caption{\label{fig:unnamed-chunk-1}Switchboard frequency-based information curves}
\end{figure}

We have found a unique average distribution of information that appears to characterize the English language as a whole, regardless of medium (spoken or written) or utterance length. This distribution indicates that in English, the words we speak or write at the beginnings of utterances have little information, while the words we speak or write at the ends of utterances have a lot of information. The words in the middle of utterances have a middling amount of information, without the increasing trend in information from word to word that we might expect from (Genzel \& Charniak, 2002).

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-2-1.pdf}
\caption{\label{fig:unnamed-chunk-2}Switchboard bigram context-based information curves}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-3-1.pdf}
\caption{\label{fig:unnamed-chunk-3}Switchboard trigram context-based information curves}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-4-1.pdf}
\caption{\label{fig:unnamed-chunk-4}Providence frequency-based information curves}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-5-1.pdf}
\caption{\label{fig:unnamed-chunk-5}Providence bigram context-based information curves}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-6-1.pdf}
\caption{\label{fig:unnamed-chunk-6}Providence trigram context-based information curves}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-7-1.pdf}
\caption{\label{fig:unnamed-chunk-7}CHILDES frequency-based information curves}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-8-1.pdf}
\caption{\label{fig:unnamed-chunk-8}CHILDES bigram context-based information curves}
\end{figure}

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-9-1.pdf}
\caption{\label{fig:unnamed-chunk-9}CHILDES trigram context-based information curves}
\end{figure}

@@

\includegraphics{paper_files/figure-latex/unnamed-chunk-10-1.pdf}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-11-1.pdf}
\caption{\label{fig:unnamed-chunk-11}Some trigram information curves from the Wikipedia data}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-12-1.pdf}
\caption{\label{fig:unnamed-chunk-12}Some linear model results from Wikipedia and WALS features}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-13-1.pdf}
\caption{\label{fig:unnamed-chunk-13}wals features vs cosine similarity}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{paper_files/figure-latex/unnamed-chunk-14-1.pdf}
\caption{\label{fig:unnamed-chunk-14}ldn features vs cosine similarity}
\end{figure}

### 2019-09-12 surprisal/bnc&switchboard.Rmd
paper citations
bootstrap childes surprisal results

# 
# sw <- switchboard %>%
#   left_join(switch_unigrams) %>%
#   mutate(s = -log(p)) %>%
#   ungroup() %>%
#   select(length, word_order, s) %>%
#   nest(s)
# 
# swc <- sw %>% filter(length %in% c(10, 20, 30))
# switch_surprisals <- map_dfr(1:nrow(swc), 
#         ~bind_cols(select(slice(swc, .x), length, word_order), 
#                                tidyboot_mean(data = unnest(slice(swc, .x)), 
#                                              column = s, nboot = 1000)))

@@

bigram_surprisals_prep <- switchboard %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  left_join(switch_bigrams) %>%
  left_join(switch_unigrams)

switch_bigram_surprisals <- bigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)

@@

trigram_surprisals_prep <- switchboard %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  left_join(switch_trigrams) %>% 
  left_join(switch_bigrams, by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
  left_join(switch_unigrams, by = c("word" = "word"))

switch_trigram_surprisals <- trigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
                    ifelse(is.na(lag_word1), # check if not the second word
                           -log(p), # trigram if third or beyond
                           -log(cond_p)), # bigram if second word
                    -log(tri_cond_p))) %>% # unigram if first word
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)

@@

>>>>

sw <- switchboard %>%
  left_join(switch_unigrams) %>%
  mutate(s = -log(p)) %>%
  ungroup() %>%
  select(length, word_order, s) %>%
  nest(s) %>%
  filter(length %in% c(7, 9, 11))

switch_surprisals <- map_dfr(1:nrow(sw),
        ~bind_cols(select(slice(sw, .x), length, word_order),
                               tidyboot_mean(data = unnest(slice(sw, .x)),
                                             column = s, nboot = 1000)))

@@

bigram_surprisals_prep <- switchboard %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  ungroup() %>%
  left_join(switch_bigrams, by = c("word", "lag_word")) %>%
  left_join(switch_unigrams, by = "word")

sbs <- bigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  ungroup() %>%
  select(length, word_order, s) %>%
  nest(s) %>%
  filter(length %in% c(7, 9, 11))

switch_bigram_surprisals <- map_dfr(1:nrow(sbs),
        ~bind_cols(select(slice(sbs, .x), length, word_order),
                               tidyboot_mean(data = unnest(slice(sbs, .x)),
                                             column = s, nboot = 1000)))

# switch_bigram_surprisals <- bigram_surprisals_prep %>%
#   mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
#   group_by(length, word_order) %>%
#   summarise(s = mean(s)) %>%
#   filter(word_order <= length)

@@

trigram_surprisals_prep <- switchboard %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  left_join(switch_trigrams) %>% 
  left_join(switch_bigrams, by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
  left_join(switch_unigrams, by = c("word"))

sts <- trigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
                    ifelse(is.na(lag_word1), # check if not the second word
                           -log(p), # trigram if third or beyond
                           -log(cond_p)), # bigram if second word
                    -log(tri_cond_p))) %>% # unigram if first word
  ungroup() %>%
  select(length, word_order, s) %>%
  nest(s) %>%
  filter(length %in% c(7, 9, 11))

switch_trigram_surprisals <- map_dfr(1:nrow(sts),
        ~bind_cols(select(slice(sts, .x), length, word_order),
                               tidyboot_mean(data = unnest(slice(sts, .x)),
                                             column = s, nboot = 1000)))
# switch_trigram_surprisals <- trigram_surprisals_prep %>%
#   mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
#                     ifelse(is.na(lag_word1), # check if not the second word
#                            -log(p), # trigram if third or beyond
#                            -log(cond_p)), # bigram if second word
#                     -log(tri_cond_p))) %>% # unigram if first word
#   group_by(length, word_order) %>%
#   summarise(s = mean(s)) %>% 
#   filter(word_order <= length)

@@

```{r saving switchboard surprisals}
switch_surprisals %>% 
  mutate(context = 0, corpus = "Switchboard") %>%
  bind_rows(switch_bigram_surprisals %>% mutate(context = 1, corpus = "Switchboard")) %>% 
  bind_rows(switch_trigram_surprisals %>% mutate(context = 2, corpus = "Switchboard")) %>% 
  write_feather(here("Data/switch_bootstrap.feather"))
```

### 2019-09-12 surprisal/childes_surprisal.Rmd
bootstrap childes surprisal results

Get utterances
```{r get_utterances}
prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
zhou_utterances <- get_utterances(corpus = "ZhouDinner") #stem for pinyin
okayama_utterances <- get_utterances(corpus = "Okayama")
```

@@

```{r hard-coded language unigrams}
child_utterances <- prov_utterances %>%
  filter(speaker_role == "Target_Child") %>% 
  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n())

child_surprisals <- child_utterances %>%
  left_join(child_unigrams) %>%
  mutate(s = -log(p)) %>%
  group_by(length, word_order) %>%
  summarise(mean = mean(s),
            median = median(s),
            se = sd(s)/sqrt(n()))

adult_utterances <- prov_utterances %>%
  filter(speaker_role != "Target_Child") %>% 
  mutate(length = str_count(gloss, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, gloss, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n())

adult_unigrams <- adult_utterances %>%
  group_by(word) %>%
  count() %>%
  ungroup() %>%
  mutate(p = n / sum(n))

adult_surprisals <- adult_utterances %>%
  left_join(adult_unigrams) %>%
  mutate(s = -log(p)) %>%
  group_by(length, word_order) %>%
  summarise(mean = mean(s),
            se = sd(s)/sqrt(n()))

@@

```{r bigrams}
adult_bigrams <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  left_join(adult_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = n.x/n.y)

bigram_surprisals_prep <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  left_join(adult_bigrams) %>%
  left_join(adult_unigrams)

bigram_surprisals <- bigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)

@@

```{r trigrams}
adult_trigrams <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  group_by(lag_word2, lag_word1, word) %>%
  count() %>%
  filter(!is.na(lag_word1), !is.na(lag_word2)) %>%
  ungroup() %>%
  left_join(adult_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
  mutate(tri_cond_p = n / n.x) %>%
  select(-n, -n.x, -n.y, -p, -cond_p)

trigram_surprisals_prep <- adult_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  left_join(adult_trigrams) %>% 
  left_join(select(adult_bigrams, lag_word, word, cond_p), by = c("lag_word1" = "lag_word", "word" = "word")) %>% 
  left_join(adult_unigrams)

trigram_surprisals <- trigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
                    ifelse(is.na(lag_word1), # check if not the second word
                           -log(p), # trigram if third or beyond
                           -log(cond_p)), # bigram if second word
                    -log(tri_cond_p))) %>% # unigram if first word
  group_by(length, word_order) %>%
  summarise(s = mean(s)) %>% 
  filter(word_order <= length)
```

>>>>

Get utterances
```{r get_utterances}
prov_utterances <- get_utterances(corpus = "Providence") #stem or gloss
zhou_utterances <- get_utterances(corpus = "ZhouDinner") #stem for pinyin
okayama_utterances <- get_utterances(corpus = "Okayama") # gloss
shiro_utterances <- get_utterances(corpus = "Shiro") #gloss
palasis_utteranecs <- get_utterances(corpus = "Palasis") #stem
wagner_utterances <- get_utterances(corpus = "Wagner") #gloss
```

@@

@@

```{r hard-coded language unigrams}
child_utterances <- palasis_utterances %>%
  filter(speaker_role != "Target_Child") %>% 
  mutate(length = str_count(stem, pattern = "[ +]+") + 1) %>%
  mutate(utterance_id = 1:n()) %>%
  unnest_tokens(word, stem, token = stringr::str_split, pattern = "[ +]+") %>%
  group_by(utterance_id) %>%
  mutate(word_order = 1:n())

child <- child_utterances %>%
  left_join(child_unigrams) %>%
  mutate(s = -log(p)) %>%
  ungroup() %>%
  select(length, word_order, s) %>%
  nest(s) %>%
  filter(length %in% c(5, 7, 9))

child_surprisals <- map_dfr(1:nrow(child),
        ~bind_cols(select(slice(child, .x), length, word_order),
                               tidyboot_mean(data = unnest(slice(child, .x)),
                                             column = s, nboot = 1000)))

@@

```{r bigrams}
child_bigrams <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  group_by(lag_word, word) %>%
  count() %>%
  filter(!is.na(lag_word)) %>%
  ungroup() %>%
  left_join(child_unigrams, by = c("lag_word" = "word")) %>%
  mutate(cond_p = n.x/n.y) %>%
  select(-n.y, -p)

bigram_surprisals_prep <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word = lag(word)) %>%
  left_join(child_bigrams, by = c("lag_word", "word")) %>%
  left_join(child_unigrams, by = "word")

child <- bigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word), -log(p), -log(cond_p))) %>%
  ungroup() %>%
  select(length, word_order, s) %>%
  nest(s) %>%
  filter(length %in% c(5, 7, 9))

child_bigram_surprisals <- map_dfr(1:nrow(child),
        ~bind_cols(select(slice(child, .x), length, word_order),
                               tidyboot_mean(data = unnest(slice(child, .x)),
                                             column = s, nboot = 1000)))

```{r trigrams}
child_trigrams <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  group_by(lag_word2, lag_word1, word) %>%
  count() %>%
  filter(!is.na(lag_word1), !is.na(lag_word2)) %>%
  ungroup() %>%
  left_join(child_bigrams, by = c("lag_word2" = "lag_word", "lag_word1" = "word")) %>%
  mutate(tri_cond_p = n / n.x) %>%
  select(-n, -n.x, -cond_p)

trigram_surprisals_prep <- child_utterances %>%
  group_by(utterance_id) %>%
  mutate(lag_word1 = lag(word)) %>%
  mutate(lag_word2 = lag(lag_word1)) %>%
  left_join(child_trigrams) %>% 
  left_join(select(child_bigrams, lag_word, word, cond_p), by = c("lag_word1" = "lag_word", "word")) %>% 
  left_join(child_unigrams, by = "word")

child <- trigram_surprisals_prep %>%
  mutate(s = ifelse(is.na(lag_word2), # check if it's not the first word
                    ifelse(is.na(lag_word1), # check if not the second word
                           -log(p), # trigram if third or beyond
                           -log(cond_p)), # bigram if second word
                    -log(tri_cond_p))) %>% # unigram if first word
  ungroup() %>%
  select(length, word_order, s) %>%
  nest(s) %>%
  filter(length %in% c(5, 7, 9))

child_trigram_surprisals <- map_dfr(1:nrow(child),
        ~bind_cols(select(slice(child, .x), length, word_order),
                               tidyboot_mean(data = unnest(slice(child, .x)),
                                             column = s, nboot = 1000)))
```

```{r bootstrap data write out}
surprisals <- surprisals %>% 
  bind_rows(child_surprisals %>% mutate(context = 0) %>%
  bind_rows(child_bigram_surprisals %>% mutate(context = 1)) %>% 
  bind_rows(child_trigram_surprisals %>% mutate(context = 2)) %>% 
  mutate(speaker = "Parent", language = "French"))
surprisals %>% 
  write_feather(here("Data/all_childes_bootstrap.feather"))
```
